\documentclass[a4paper,12pt]{article}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{geometry}
\usepackage{longtable}

% Page Layout
\geometry{margin=1in}

% Code Styling
\lstset{
	basicstyle=\ttfamily\footnotesize,
	keywordstyle=\color{blue}\bfseries,
	stringstyle=\color{green!60!black},
	commentstyle=\color{gray},
	numbers=left,
	numberstyle=\tiny\color{gray},
	stepnumber=1,
	numbersep=10pt,
	backgroundcolor=\color{gray!10},
	frame=single,
	breaklines=true,
	tabsize=4,
	captionpos=b
}

\title{Overview of C/C++ Compilers}
\author{Deva Nayak}
\date{\today}

\begin{document}

\maketitle

\section{Introduction to C/C++ Compilers}
A compiler translates human-readable C/C++ source code into machine-readable executable code. Compilers are essential for creating applications and are specific to platforms, operating systems, and sometimes hardware.

\subsection{Key Features of a Compiler}
\begin{itemize}
	\item Converts source code into machine code.
	\item Ensures platform-specific compatibility.
	\item Provides debugging, optimization, and error diagnostics.
\end{itemize}

---

\section{Compilers by Operating Systems}

\subsection{Linux}
\begin{itemize}
\item \textbf{GCC (GNU Compiler Collection):}
\begin{itemize}
	\item Open-source and widely used for C and C++.
	\item Rich debugging and optimization options.
\end{itemize}
Example Command:
\begin{lstlisting}[language=bash]
gcc -o program program.c
g++ -o program program.cpp
\end{lstlisting}

\item \textbf{Clang (LLVM):}
\begin{itemize}
	\item Modern, fast, and modular compiler.
	\item Compatible with GCC commands.
\end{itemize}
Example Command:
\begin{lstlisting}[language=bash]
clang -o program program.c
clang++ -o program program.cpp
\end{lstlisting}
\end{itemize}

\subsection{Windows}
\begin{itemize}
\item \textbf{MinGW (Minimalist GNU for Windows):}
\begin{itemize}
	\item GCC-based compiler for Windows.
	\item Lightweight and straightforward.
\end{itemize}
Example Command:
\begin{lstlisting}[language=bash]
gcc -o program.exe program.c
\end{lstlisting}

\item \textbf{Microsoft Visual C++ (MSVC):}
\begin{itemize}
	\item Part of Visual Studio.
	\item Provides advanced debugging tools and Windows optimization.
\end{itemize}
Example Command:
\begin{lstlisting}[language=bash]
cl /EHsc program.cpp
\end{lstlisting}

\item \textbf{Cygwin:}
\begin{itemize}
	\item Unix-like environment for Windows with GCC support.
\end{itemize}
\end{itemize}

\subsection{macOS}
\begin{itemize}
\item \textbf{Clang (Default):}
\begin{itemize}
	\item Comes with Xcode tools.
	\item Optimized for macOS.
\end{itemize}
Example Command:
\begin{lstlisting}[language=bash]
clang -o program program.c
clang++ -o program program.cpp
\end{lstlisting}

\item \textbf{GCC:}
\begin{itemize}
	\item Available via Homebrew.
\end{itemize}
Example Command:
\begin{lstlisting}[language=bash]
brew install gcc
gcc-<version> -o program program.c
\end{lstlisting}
\end{itemize}

---

\section{Compilers for Embedded Targets}

\subsection{ARM (Embedded Systems)}
\begin{itemize}
\item \textbf{ARM GCC (GNU Arm Embedded Toolchain):}
\begin{itemize}
	\item Free, open-source compiler for ARM Cortex-M boards.
\end{itemize}
Example Command:
\begin{lstlisting}[language=bash]
arm-none-eabi-gcc -o program.elf program.c
\end{lstlisting}

\item \textbf{Keil MDK:}
\begin{itemize}
	\item Proprietary IDE and compiler for ARM devices.
\end{itemize}

\item \textbf{IAR Embedded Workbench:}
\begin{itemize}
	\item Commercial compiler for embedded systems with advanced optimization.
\end{itemize}
\end{itemize}

\subsection{AVR Microcontrollers}
\begin{itemize}
\item \textbf{AVR-GCC:}
\begin{itemize}
	\item GCC-based compiler for AVR chips (e.g., Arduino).
\end{itemize}
Example Command:
\begin{lstlisting}[language=bash]
avr-gcc -o program program.c
\end{lstlisting}
\end{itemize}

\subsection{ESP32/ESP8266}
\begin{itemize}
	\item \textbf{Xtensa GCC Toolchain:}
	\begin{itemize}
		\item GCC-based toolchain for ESP32/ESP8266.
		\item Works with ESP-IDF (Espressif IoT Development Framework).
	\end{itemize}
\end{itemize}

\subsection{RISC-V}
\begin{itemize}
	\item \textbf{RISC-V GCC Toolchain:}
	\begin{itemize}
		\item GCC-based toolchain for RISC-V processors.
	\end{itemize}
\end{itemize}

---

\section{Summary of Popular Compilers}

\begin{tabular}{|p{0.18\linewidth}|p{0.18\linewidth}|p{0.30\linewidth}|p{0.26\linewidth}|}
	\hline
	\textbf{Compiler} & \textbf{OS} & \textbf{Features} & \textbf{Usage} \\ \hline
	GCC & Linux, macOS, MinGW & Open-source, cross-platform & General-purpose \\ \hline
	Clang & macOS, Linux, Windows & Modern, modular, great diagnostics & Performance-critical projects \\ \hline
	MSVC & Windows & Microsoft-supported, Windows optimization & Windows applications \\ \hline
	ARM GCC & Cross-platform & Free toolchain for ARM microcontrollers & Embedded systems \\ \hline
	Keil MDK & Windows & Proprietary, ARM Cortex-M optimization & Professional embedded projects \\ \hline
	IAR Embedded Workbench & Cross-platform & Highly optimized & Embedded development \\ \hline
	AVR-GCC & Cross-platform & Open-source for AVR microcontrollers & Arduino projects \\ \hline
	Xtensa GCC & Cross-platform & Toolchain for ESP32/ESP8266 & IoT development \\ \hline
\end{tabular}
\end{document}
